---
title: "Annex III.7 - STREAM: Auxiliary scripts for the conversion from the SDEF format to DG MARE Med&BS Data Call: Calculation of sex ratio at age using the COST methods (see COST User Manual- Chapter 12, for the full method description)"
author: "E. Mantzouni"
date: "`r date()`"
output: word_document
---



#Tools

R, Rstudio and packages.


```{r results='hide', message=FALSE, warning=FALSE}
#R general option:
options(stringsAsFactors = FALSE)

options(warn=0)

#chunk option
knitr::opts_chunk$set(cache=TRUE,echo=TRUE, warning=FALSE, 
	message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
		include=TRUE,dev='png',autodep=FALSE)



#load packages

library(COSTcore)
library(COSTeda)
library(COSTdbe)

library(tidyr)
library(dplyr)
library(knitr)
```


#Example of use of the script SRA

This script implements the  Calculation of  sex ratio at age, 
required for the DG MARE Med&BS Data Call, using as input file the SDEF format (CS table) and COST as the raising proceedure


##Settings 

```{r data1}

path <- paste("C:\\Users\\Bitetto Isabella\\OneDrive - Coispa Tecnologia & Ricerca S.C.A.R.L\\MARE22\\STREAM\\FINAL REVISION OF DELIVERABLES\\SDEF_to_DG_MARE_MedBS", sep="") 

setwd(path)
  
```

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = path)
```

##Input Data

csData of the SDEF format 

```{r data_in1}

load("fri_cs_test.Rdata")

head( fri_cs@ca)


# extract COUNTRY
COUNTRY<-unique(fri_cs@ca$landCtry)

```
 stratification: spatial only
```{r data_str}


fri_strD <- strIni(spaceStrata="area") 

```

DG MARE Med&BS SRA Table template

```{r data_in2}

sra.temp2<- read.table("xxx_SRA.csv",sep=",",header=T)

names(sra.temp2)
```


Auxilliary table: species_BIO_SRA.csv


```{r data_in3}

# species file : selected species with FAO three alpha code

sel_spe<-read.table( "species_BIO_SRA.csv", header=TRUE, sep=";",row.names=NULL) 

head(sel_spe)

```


## Data analysis- raising

Analysis by stock
```{r loop by stock}

for (i in 1:dim(sel_spe)[1]) {

  STK<- sel_spe$SPECIES[i]
  
    fri_cs1<- subset(fri_cs, year%in% seq(sel_spe$START_YEAR[i],
                                          sel_spe$END_YEAR[i],by=1),table="ca",link=T)
   
# Validate- consolidate 
fri_csv <- csDataVal(fri_cs1)
fri_csc<- csDataCons(fri_csv, fri_strD)
  
  fri_csc1<- subset(fri_csc, space==sel_spe$GSA[i],table="ca")
  
MS_SEX_An<-dbeObject(desc="SEX-ratio",species=STK,param="sex",
                     strataDesc=fri_strD,methodDesc=sel_spe$methodDesc[i],
                     nboot=sel_spe$nboot[i])


if (sel_spe$methodDesc[i]=="bootstrap"){
MS_SEX_An<-bpBoot(MS_SEX_An,fri_csc1,adjust=sel_spe$adjust[i])
} else {
MS_SEX_An<-bpEstim(MS_SEX_An,fri_csc1,adjust=sel_spe$adjust[i])
}

dfSRA <-
  data.frame(
    
    COUNTRY = COUNTRY ,
    AREA = MS_SEX_An@ageStruc[["estim"]][["space"]] ,
     START_YEAR = sel_spe$START_YEAR[i] ,
    END_YEAR = sel_spe$END_YEAR[i] ,
    SPECIES = STK ,
    AGECLASS =  MS_SEX_An@ageStruc[["estim"]][["age"]] ,
    SEX_RATIO = MS_SEX_An@ageStruc[["estim"]][["value"]],
    COMMENTS = sel_spe$COMMENTS[i] 
  )


dfSRA<-dfSRA %>% mutate(SPECIES=sel_spe$SPE[match(STK,sel_spe$SPECIES)])

dfSRA <- dfSRA[complete.cases(dfSRA[, -which(names(dfSRA) %in% c("COMMENTS"))]),]

dfSRA$AGECLASS <- as.numeric(as.character(dfSRA$AGECLASS))

sra.temp2=bind_rows(sra.temp2, dfSRA)


}



```

##Output

```{r data_ou1}

# export DG MARE Med&BS SRA table
write.table(format(sra.temp2,digits=3, scientific=F),
            file="SRA.csv",dec=".",sep=",",col.names=TRUE,
            row.names=FALSE,na="-1")

head(sra.temp2)
```



