---
title: "Annex III.8 - STREAM: Auxiliary scripts for the conversion from the SDEF format to DG MARE Med&BS Data Call: Calculation of Age Length Key (ALK) using the COST methods (see COST User Manual- Chapter 10, for the full method description)"
author: "E. Mantzouni"
date: "`r date()`"
output: word_document
---



#Tools

R, Rstudio and packages.


```{r results='hide', message=FALSE, warning=FALSE}
#R general option:
options(stringsAsFactors = FALSE)

options(scipen = 999) # disable scientific notation

options(warn=0)

#chunk option
knitr::opts_chunk$set(cache=TRUE,echo=TRUE, warning=FALSE, 
                      message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
                      include=TRUE,dev='png',autodep=FALSE)



#load packages

library(COSTcore)
library(COSTeda)
library(COSTdbe)

library(tidyr)
library(dplyr)
library(data.table)
```


#Example of use of the script ALK

This script implements the  Calculation of  the age length key, 
required for the DG MARE Med&BS Data Call, using as input file the SDEF format (CS table) and COST as the raising procedure
Additional methods exist for filling in the potential gaps for length classes 


##Settings 

```{r data1}

path <- paste("C:\\Users\\Bitetto Isabella\\OneDrive - Coispa Tecnologia & Ricerca S.C.A.R.L\\MARE22\\STREAM\\FINAL REVISION OF DELIVERABLES\\SDEF_to_DG_MARE_MedBS", sep="") 

setwd(path)



```
```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = path)
```

##Input Data

csData of the SDEF format 

```{r data_in1,warning=FALSE}
load("fri_cs_test.Rdata")
head( fri_cs@ca)

# extract COUNTRY
COUNTRY<-unique(fri_cs@ca$landCtry)

```

stratification: spatial only
```{r data_str}
fri_strD <- strIni(spaceStrata="area") 
```

DG MARE Med&BS ALK Table template

```{r data_in2}

alk.temp2<- read.table("xxx_ALK.csv",sep=",",header=T)
head(alk.temp2)
```


Auxiliary table: species_BIO_ALK.csv


```{r data_in3, echo=TRUE, message=TRUE, warning=FALSE}

# species file : selected species with FAO three alpha code 

sel_spe<-read.table( "species_BIO_ALK.csv", header=TRUE, sep=";",row.names=NULL) 

head(sel_spe)

sel_spe<- sel_spe %>% 
  mutate(SEX = strsplit(as.character(SEX), "_")) %>% 
  unnest(SEX)

```


## Data analysis- raising

Analysis by stock
```{r}


for (i in 1:dim(sel_spe)[1]) {
  
  STK<- sel_spe$SPECIES[i]
  
  fri_cs1<- subset(fri_cs, year%in% seq(sel_spe$START_YEAR[i],
                    sel_spe$END_YEAR[i],by=1),table="ca",link=T)
  
  #  estimate sample size (number of otoliths per stock, sex and age)  
  if (sel_spe$SEX[i]=="C"){
    
    fri_csv <- csDataVal(fri_cs1)
  
    nml<- data.frame(fri_cs1@ca) %>% filter(!is.na(age))%>% 
      dplyr::group_by(area,spp,age)%>% 
      summarize(TOTAL_NUMBER_OF_HARD_STRUCTURE_READ_BY_AGE=n())
    
  } else { # ALK for selected sex
    
    fri_cs1=subset(fri_cs1,sex==sel_spe$SEX[i],table="ca",link=T)
    fri_csv <- csDataVal(fri_cs1)
    
    # get sample size: number of otoliths
    nml<- data.frame(fri_cs1@ca) %>% filter(!is.na(age))%>% 
      dplyr::group_by(area,spp,age,sex)%>% 
      summarize(TOTAL_NUMBER_OF_HARD_STRUCTURE_READ_BY_AGE=n())
    
  }
  
  
  fri_csv1<- subSetSpp(fri_csv, STK)
  fri_csv1<- subset(fri_csv1, area%in% sel_spe$GSA[i],table="ca",link=T)
  fri_csc1 <- csDataCons(fri_csv1, fri_strD)
  
  ## ### CV from individual length-at-age 
LEstim_An <-
dbeObject(
desc = "Length at age",
species = STK,
catchCat = "LAN",
param = "length",
strataDesc = fri_strD, # ,
methodDesc = "analytical"   #sel_spe$methodDesc_LAN.len.age[i]
)


if(sel_spe$methodDesc_LAN.len.age[i]=="analytical") {
LEstim_An <-  bpEstim(LEstim_An, fri_csc1, adjust = sel_spe$adjust[i])

} else{
LEstim_An <-  bpBoot(LEstim_An, fri_csc1, adjust = sel_spe$adjust[i])

}


## ALK
  
  res1 <- alkLgthRec(fri_csc1,type=sel_spe$typeALK[i],value=sel_spe$valueALK[i],
                     update=F, preview=F,postview = F)
  
  if (sel_spe$typeALK[i]=="fillALKmult"){
    fri_csc2 <- fillALKmult(fri_csc1,STK,p=10,trace=T)
    res1 <- alkLgthRec(fri_csc2,update=F, preview=F,postview = F,
                       value=sel_spe$valueALK[i])
  }
  
  dfALK <-
    data.frame(
      COUNTRY =COUNTRY ,
      AREA =sel_spe$GSA[i],
      START_YEAR = unique(sel_spe$START_YEAR[i]) ,
      END_YEAR = unique( sel_spe$END_YEAR[i]) ,
      SPECIES = STK ,
      SEX = sel_spe$SEX[i],
      UNIT = unique(fri_cs1@ca$lenCode[fri_cs1@ca$spp==STK]) ,
      SPECON=  -1,
      APPLY_TO_CATCHES_FILE= unique( sel_spe$APPLY_TO_CATCHES_FILE[i]),
      CV=NA,
      AGE = as.numeric(colnames(res1$alk)),
      COMMENTS= sel_spe$COMMENTS[i],
         LENGTHCLASS100_PLUS=0
    )
  
  # get sample size
  
  if (sel_spe$SEX[i]=="C"){
    
    dfALK <-  dfALK %>%
      left_join(nml, by = c("AREA" = 'area', 'SPECIES' = 'spp',"AGE"="age")) %>% 
      mutate(SPECIES =  sel_spe$SPE[i]) 
    # FAO Three alpha code
  } else{
    
    dfALK <-  dfALK %>%
      left_join(nml, by = c("AREA" = 'area', 'SPECIES' = 'spp',"AGE"="age", 
                            "SEX"="sex")) %>% 
      mutate(SPECIES =  sel_spe$SPE[i]) 
    # FAO Three alpha code
    
  }
  
  
  aa=data.frame(res1$alk)
  names(aa)=colnames(res1$alk)
  
    aa=aa%>% mutate(LC=rownames(res1$alk))
    
    ## fix LC 
   UNIT <- as.character( unique(fri_csc1@ca$lenCode[fri_csc1@ca$spp==STK]) )
  
 if (UNIT %in% c("mm", "MM")& sel_spe$LC_RANGE[i]==10) {
 aa$LC<-as.numeric(aa$LC)/10
 UNIT1<-"cm"
 }
  
  if (UNIT %in% c("mm", "MM") & sel_spe$LC_RANGE[i]==1) {
 aa$LC<-as.numeric(aa$LC)
 UNIT1<- "mm"
  }
 
   if (UNIT %in% c("mm", "MM")& sel_spe$LC_RANGE[i]==5) {
 aa$LC<-as.numeric(aa$LC)/10
 UNIT1<-"cm"
  }
 
   if (UNIT %in% c("cm", "CM") ) {
 aa$LC<-as.numeric(aa$LC)
 UNIT1<- "cm"
  }
 
   aa$LC<- plyr::round_any( aa$LC,1,floor)
###
   
  dfALK$UNIT<-UNIT1 
  
  aa1=aa %>%  gather(age, n.at.len, -LC)
  
  aa1 <- as.data.table(aa1)

  seq_l <-  seq(0, max(aa$LC,na.rm = T), by = 1) # 
  dt1<- aa1[, list(LC = seq_l), by = age]
  
  
  dt2<- left_join(dt1,aa1)

   dt3 <- data.table::dcast(dt2,as.formula(paste(paste(names(dt2)[! names(dt2)
                      %in% c("LC","n.at.len")], collapse='+'), "LC", sep="~")), 
                      value.var = "n.at.len")
  
   dt3$age=as.numeric(dt3$age)
  
  
  dt3<- dt3 %>% mutate_at(vars( -(age) ), 
          funs( if_else( is.na(.), 0, .) ) )
  
  dfALK<-left_join(dfALK,dt3,by=c("AGE"="age"))
  
  ## CV
 # LEstim_An @ageNum$ cv
 
  LEstim_An@ageNum[["cv"]]$age=as.numeric(LEstim_An@ageNum[["cv"]]$age)
    
   dfALK<- dfALK%>% left_join(LEstim_An@ageNum[["cv"]]%>% select(age,value),
                by=c( "AGE"="age"))%>% mutate(CV=value)%>% select(-c(value))

  # take care of number of Length classes (max is 100 acc. to DG MARE Med&BS template)
  zz<-dim(dfALK[-c(1:14)])[2]
  names(dfALK)[-c(1:14)]<- paste("LENGTHCLASS",seq(0,zz-1,1),sep="")
  
  
  if(zz>=100){
    dfALK$LENGTHCLASS100_PLUS<- rowSums(dfALK[,-c(1:114)])
  }
  
   dfALK<-dfALK%>% select(one_of(as.vector(names(alk.temp2))))

  alk.temp2<- bind_rows(alk.temp2,(dfALK))
  
 alk.temp2[,-c(1:13,114)][is.na(alk.temp2[,-c(1:13,114)])] <- 0
  
  
  # export updated CsDataCons
  
  if (sel_spe$APPLY_TO_CATCHES_FILE[i] == "Y") {
    if (sel_spe$typeALK[i] == "fillALKmult") {
      fri_csc2 <- fillALKmult(fri_csc1, STK, p = 10, trace = T)
      
      save(fri_csc2,
           file = paste("upd", STK, sel_spe$SEX[i], sel_spe$GSA[i], ".Rdata", sep ="_"))
      
    } else{
      res1 <-
        alkLgthRec(
          fri_csc1,
          type = sel_spe$typeALK[i],
          value = sel_spe$valueALK[i],
          update = T,
          preview = F,
          postview = F
        )
      
      save(res1,
           file = paste("upd", STK, sel_spe$SEX[i], sel_spe$GSA[i], ".Rdata", sep =
                          "_"))
    }
    
  }
  
}



```

##Output

```{r data_ou1}

# export DG MARE Med&BS ALK table
write.table(format(alk.temp2,digits=3, scientific=F),
    file="ALK.csv",dec=".",sep=",",col.names=TRUE,
    row.names=FALSE,na="-1")

head(alk.temp2)


```



