---
title: "Annex III.1 - STREAM: Auxiliary scripts for the conversion from the SDEF format to DG MARE Med&BS Data Call: Raising of the length distributions by ID to the total landings"
author: "E. Mantzouni"
date: "`r date()`"
output: word_document
---

```{r results='hide', message=FALSE, warning=FALSE}
#R general option:
options(stringsAsFactors = FALSE)

options(warn=0)

options(scipen = 999) # disable scientific notation

#chunk option
knitr::opts_chunk$set(cache=TRUE,echo=TRUE, warning=FALSE, 
	message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
		include=TRUE,dev='png',autodep=FALSE)

#load packages

library(COSTcore)
library(COSTeda)
library(COSTdbe)

library(tidyr)
library(dplyr)
library(data.table)
library(knitr)

```




#Example of use of the script B_LANDINGS

This script implements the Calculation of the raised numbers by length required for the DG MARE Med&BS Data Call, using as input file the SDEF format (CS table) and COST as the raising procedure

##Settings 

```{r data1  , message=FALSE}

path_in <- paste("C:\\Users\\Bitetto Isabella\\OneDrive - Coispa Tecnologia & Ricerca S.C.A.R.L\\MARE22\\STREAM\\FINAL REVISION OF DELIVERABLES\\SDEF_to_DG_MARE_MedBS", sep="") 

setwd(path_in)

```
```{r "setup", include=FALSE}

opts_knit$set(root.dir = path_in)
```

##Input Data

Validated csData and clData of the SDEF format and stratification object

```{r data_in1,warning=FALSE}
load("fri-TEST.Rdata")


fri_strD1 <- strIni(timeStrata="quarter", techStrata = "foCatEu5",
                    spaceStrata = "area")


fri_strD1


  fri_csc <- csDataCons(fri_cs1, fri_strD1)

fri_clc <- clDataCons(fri_cl1, fri_strD1)

head(fri_csc )
head(fri_clc)

# extract COUNTRY  and YEAR
COUNTRY<-unique(fri_cl1@cl$landCtry)
YEAR=unique(fri_cl1@cl$year)

```


DG MARE Med&BS LANDINGS Table template

```{r data_in3}

lan.temp2<- read.table("xxx_LANDINGS.csv",sep=",",header=T)

head(lan.temp2)
```

Communication Table for FISHERY

```{r data_in4}

fishery<- read.table("communicationTable_for_fishery.csv",sep=";",header=T)

head(fishery)
```

Auxiliary table:  species_LAN.csv


```{r data_in5, echo=TRUE, message=TRUE, warning=FALSE}

# species file : selected species with FAO three alpha code

sel_spe <-read.table("species_LANDINGS.csv",sep=";",header=T)

head(sel_spe)

```

## Data analysis- raising

Analysis by stock


```{r loop by stock, warning=F, message=F}




for (i in 1:dim(sel_spe)[1]) {
  
  STK<- sel_spe$SPECIES[i]
  
  AREA <- sel_spe$GSA[i]
  
    fri_csc1<- subset(fri_csc, space==sel_spe$GSA[i],table="ca",link=T)
  fri_clc1<- subset(fri_clc, space==sel_spe$GSA[i],table="cl")

  # The first step is to create the empty object, that will be given 
  # the appropriate values for the descritor  fields.
  
  lanEstim <-
    dbeObject(
      desc = paste(STK, AREA,"Landings", sep="_"),
      species = STK,
      catchCat = "LAN",
      strataDesc = fri_strD1,
      methodDesc = sel_spe$lanEstim_methodDesc[i]
    )
  
  # the only arguments to pass to the function are the dbe object, 
  # the consolidated cs and cl datasets.

      if ( sel_spe$lanEstim_methodDesc[i]=="analytical"){
lanEstim <- RaiseLgth(lanEstim, fri_csc1, fri_clc1,incl.precision =F)
     } else {
  lanEstim <- RaiseLgthBoot(lanEstim, fri_csc1, fri_clc1,
                            incl.precision =F,B=15)
} 
  

  # totalW\$estim : total weight, 
aa <-lanEstim@totalW$estim 
  
aa$value<- aa$value/1000 # tons

aa<- rename(aa, "totalW"=value)
  
# lenStruc\$estim : numbers-at-length estimates 

  bb<- lanEstim@lenStruc$estim
  
   # define LCs and UNIT len 
 UNIT <- as.character( unique(fri_csc@ca$lenCode[fri_csc@ca$spp==STK]) )
  
 if (UNIT %in% c("mm", "MM")& sel_spe$LC_RANGE[i]==10) {
 bb$length<-as.numeric(bb$length)/10
 UNIT1<-"CM"
 }
  
  if (UNIT %in% c("mm", "MM") & sel_spe$LC_RANGE[i]==1) {
 bb$length<-as.numeric(bb$length)
 UNIT1<- "MM"
  }
 
   if (UNIT %in% c("mm", "MM")& sel_spe$LC_RANGE[i]==5) {
 bb$length<-as.numeric(bb$length)/10
 UNIT1<-"CM"
  }
 
   if (UNIT %in% c("cm", "CM") ) {
 bb$length<-as.numeric(bb$length)
 UNIT1<- "CM"
  }
 
   bb$length<- plyr::round_any( bb$length,1,floor)
 
  bb$value<- bb$value/1000 # '000 ind
  
  ab=left_join(bb,aa ,by = c("time", "space", "technical"))
  
  ab<- ab %>% separate(technical, c("gear","FISHERY", "VL","MESH_SIZE_RANGE"), 
                       sep = "_",remove=T)
  
  ab$length<- as.numeric(as.character(ab$length))
  
  ab<- ab%>% group_by(time,   space , gear  ,FISHERY, VL,MESH_SIZE_RANGE  ) %>%
    mutate(minlc=min(length,na.rm=T),maxlc=max(length,na.rm=T))

    # matrix with all combinations of "time"   "space"  "gear"   "VL"     
  # "length" ,"MESH_SIZE_RANGE"
 
  dt <- as.data.table(ab)
  
   dt[,c(1:7)][is.na(dt[,c(1:7)])]<- -1
  
 seq_l <-  seq(0, max(dt$length,na.rm = T), by = 1) # 
  
  dt$id<- paste(dt$time,dt$space,dt$gear,dt$FISHERY,dt$VL,
                dt$MESH_SIZE_RANGE,sep=":")
  
  dt1<- dt[, list(length = seq_l), by = id]
  
  dt1<- dt1 %>% separate(id, c("time", "space", "gear", "FISHERY","VL",
                               "MESH_SIZE_RANGE"), sep = ":")
  
 
  ab[is.na(ab)]<- -1
  
  dt2<- left_join(dt1,ab)
  dt2$stock<- STK
  
  ##
  
  dt3 <- data.table::dcast(dt2,as.formula(paste(paste(names(dt2)[! names(dt2) %in% 
                    c("length","value")], collapse='+'), "length", sep="~")), 
                    value.var = "value")
  
  
  dt3=dt3[complete.cases(dt3[,c(7:9)]), ]
  
dt3 <- dt3 %>% separate(time, c("Year","Quarter")," - ")

  dt3$MESH_SIZE_RANGE<-as.character(dt3$MESH_SIZE_RANGE)


  # numbers at LC : NA-->0  
dt3<- dt3 %>% mutate_at(vars( -(Year:stock) ), 
          funs( if_else( is.na(.), 0, .) ) )


  LANDINGS <- data.frame(
    
    ID = NA , 
    COUNTRY = COUNTRY , 
    YEAR = YEAR , 
    QUARTER =dt3$Quarter, 
    VESSEL_LENGTH = dt3$VL,
    GEAR = dt3$gear,
    MESH_SIZE_RANGE = dt3$MESH_SIZE_RANGE,
    FISHERY =  dt3$FISHERY , 
    AREA = sel_spe$GSA[i],
    SPECON = -1 , 
    SPECIES = STK , 
    LANDINGS = dt3$totalW ,
    UNIT = UNIT1
  )
  


  
  LANDINGS<-left_join(LANDINGS,dt3[,-c(1,3,8:11)],by=c( "QUARTER"  ="Quarter" ,  
                                  "GEAR"="gear" ,  "VESSEL_LENGTH" = "VL"  ,
                                  "MESH_SIZE_RANGE","FISHERY" ))
  
  # take care of number of Length classes (max is 100 acc. to DG MARE Med&BS template)
  zz<-dim(LANDINGS[-c(1:13)])[2]
  names(LANDINGS)[-c(1:13)]<- paste("LENGTHCLASS",seq(0,zz-1,1),sep="")
  

  if(zz>=100){
    
    LANDINGS$LENGTHCLASS100_PLUS<- rowSums(LANDINGS[,!1:113])
    
   LANDINGS<-LANDINGS %>% select(ID:LENGTHCLASS99,LENGTHCLASS100_PLUS)
  }
  
   

    # FISHERY to DG MARE Med&BS codification
  LANDINGS$FISHERY <- fishery$SDEF_codification[match(LANDINGS$FISHERY , 
                                    fishery$DGMARE_Med_BS_codification)]
  
  # species to FAO three alpha code and set ID (COUNTRY, AREA, GEAR, VESSEL_LENGTH, 
  # MESH_SIZE_RANGE,QUARTER, SPECIES)
  
  land.tab <-LANDINGS %>% mutate(SPECIES=sel_spe$SPE[match(SPECIES,sel_spe$SPECIES)],
                                  ID = paste(COUNTRY, AREA, GEAR,FISHERY, VESSEL_LENGTH, 
                            MESH_SIZE_RANGE,YEAR, QUARTER, SPECIES, sep = "_"))

  lan.temp2<-bind_rows(lan.temp2,land.tab)
  

lan.temp2[,-c(1:13)][is.na(lan.temp2[,-c(1:13)])] <- 0
    
}

```

##Output

```{r data_ou1}
lan.temp2 <- lan.temp2[, 2:ncol(lan.temp2)]
write.table(format(lan.temp2,digits=3,scientific=F), file = "LANDINGS.csv",
            row.names=FALSE,sep=";",na="-1")

head(lan.temp2)

  
```

