---
title: "Annex III.5 - STREAM: Auxiliary scripts for the conversion from the SDEF format to DG MARE Med&BS Data Call: Calculation of maturity at age using the COST methods (see COST User Manual- Chapter 12, for the full method description)"
author: "E. Mantzouni"
date: "`r date()`"
output: word_document
---



#Tools

R, Rstudio and packages.


```{r results='hide', message=FALSE, warning=FALSE}
#R general option:
options(stringsAsFactors = FALSE)

options(warn=0)

#chunk option
knitr::opts_chunk$set(cache=TRUE,echo=TRUE, warning=FALSE, 
	message=FALSE, fig.height=6,progress=FALSE,verbose=FALSE,
		include=TRUE,dev='png',autodep=FALSE)

#load packages

library(COSTcore)
library(COSTeda)
library(COSTdbe)

library(tidyr)
library(dplyr)
```


#Example of use of the script MA

This script implements the  Calculation of  maturity at age, 
required for the JRC  MED &BS Data Call, using as input file the SDEF format (CS table) and COST as the raising proceedure


##Settings

```{r data1}
path <- paste("C:\\Users\\Bitetto Isabella\\OneDrive - Coispa Tecnologia & Ricerca S.C.A.R.L\\MARE22\\STREAM\\FINAL REVISION OF DELIVERABLES\\SDEF_to_DG_MARE_MedBS", sep="") 
setwd(path)
  
```

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = path)
```

##Input Data

csData of the SDEF format and estimate sample size

```{r data_in1}
load("fri_cs_test.Rdata")
head( fri_cs@ca)

# extract COUNTRY
COUNTRY<-unique(fri_cs@ca$landCtry)

```
 stratification: spatial only
```{r data_str}

fri_strD <- strIni(spaceStrata="area") 

```

DG MARE Med&BS MA Table template

```{r data_in2}


ma.temp2<- read.table("xxx_MA.csv",sep=",",header=T)
names(ma.temp2)
```


Auxiliary table:  species_BIO_MA.csv

```{r data_in3}

# species file : selected species with FAO three alpha code

sel_spe<-read.table( "species_BIO_MA.csv", header=TRUE, sep=";",row.names=NULL) 

head(sel_spe)

sel_spe<- sel_spe %>% 
    mutate(SEX = strsplit(as.character(SEX), "_")) %>% 
    unnest(SEX)

imm=sel_spe%>%separate(IMMATURE,c("a", "b","c","d"), extra = "drop", 
                       fill = "right",sep="_")


```



## Checking data


Set immature stages  based on Auxiliary Table 
```{r data_immature}


# get col w. all=NA by spp
aa<- fri_cs@ca %>% group_by(spp) %>% summarize_all(~all(is.na(.))) 


# if no matScale is spcified, impute "-1"
fri_cs@ca$matScale[fri_cs@ca$spp %in% aa$spp[aa$matScale==TRUE]]<- "-1"

```

## Data analysis- raising

Analysis by stock
```{r loop by stock}

for (i in 1:dim(sel_spe)[1]) {
  
 
 STK<- sel_spe$SPECIES[i]
 
 fri_cs1<- subset(fri_cs, year%in% seq(sel_spe$START_YEAR[i],
                      sel_spe$END_YEAR[i],by=1),table="ca",link=T)
  
    
       # set immature stages as "I" 
if (na.omit(unique(fri_cs1@ca$matScale[fri_cs1@ca$spp==STK & 
              fri_cs1@ca$area==sel_spe$GSA[i]]))!="1-7") {
  fri_cs1@ca$matStage[fri_cs1@ca$spp==STK & fri_cs1@ca$area==sel_spe$GSA[i] & 
                    fri_cs1@ca$matStage %in% imm[i,6:9][!is.na(imm[i, 6:9])]]="I"
  
} else{ #if "1-7" matScale is used, immature (I)==1
  fri_cs1@ca$matStage[fri_cs1@ca$spp==STK & fri_cs1@ca$area==sel_spe$GSA[i] & 
                        fri_cs1@ca$matStage ==1]="I"
  
}
   
# sample size
  
if (sel_spe$SEX[i]=="C"){
  
    fri_csv <- csDataVal(fri_cs1)
  
    # get sample size
    
    nml<- data.frame(fri_cs1@ca) %>% filter(!is.na(matStage)& !is.na(age))%>% 
      dplyr::group_by(area,spp,age)%>% summarize(SAMPLE_SIZE=n())
    

} else { 
   
  
   fri_cs1=subset(fri_cs1,sex==sel_spe$SEX[i],table="ca",link=T)
  fri_csv <- csDataVal(fri_cs1)
  
  # get sample size
  nml<- data.frame(fri_cs1@ca) %>% filter(!is.na(matStage)& !is.na(age))%>% 
    dplyr::group_by(area,spp,age,sex)%>% summarize(SAMPLE_SIZE=n())
 
}
  
  
     fri_csv1<- subSetSpp(fri_csv, STK)
   fri_csv1<- subset(fri_csv1, area%in% sel_spe$GSA[i],table="ca")
       
  fri_csc1 <- csDataCons(fri_csv1, fri_strD)
  
  fri_csc1@ca$matStage=as.character(fri_csc1@ca$matStage)


MS_MAT_An<-dbeObject(desc=" Maturity-at-age",species=STK,param="maturity",
                     strataDesc=fri_strD,methodDesc=sel_spe$methodDesc[i],
                     nboot=sel_spe$nboot[i])


if (sel_spe$methodDesc[i]=="bootstrap"){
MS_MAT_An<-bpBoot(MS_MAT_An,fri_csc1,adjust=sel_spe$adjust[i],immature.scale="I")

} else { # analytical
MS_MAT_An<-bpEstim(MS_MAT_An,fri_csc1,adjust=sel_spe$adjust[i],immature.scale="I")
}



dfMA <-
  data.frame(
    
    COUNTRY =COUNTRY ,
    AREA = unique(MS_MAT_An@ageStruc[["estim"]][["space"]] ),
    START_YEAR = sel_spe$START_YEAR[i] ,
    END_YEAR = sel_spe$END_YEAR[i] ,
    SPECIES = STK ,
    SEX = sel_spe$SEX[i],
    
    AGECLASS =  as.numeric(as.character(MS_MAT_An@ageStruc[["estim"]][["age"]] )),

    PRM = MS_MAT_An@ageStruc[["estim"]][["value"]] ,
    METHOD_USED = sel_spe$MAT_METHOD[i]
  
  )

if (sel_spe$SEX[i]=="C"){
  
dfMA <-  dfMA %>%
           left_join(nml, by = c("AREA" = 'area', 'SPECIES' = 'spp',
                           "AGECLASS"="age")) %>%   mutate(SPECIES =  sel_spe$SPE[i]) 
                           # FAO three alpha code 
} else{
  
  dfMA <-  dfMA %>%
           left_join(nml, by = c("AREA" = 'area', 'SPECIES' = 'spp', 
                            "SEX"="sex","AGECLASS"="age")) %>% 
                            mutate(SPECIES =  sel_spe$SPE[i])
                             # FAO three alpha code 
  
}

ma.temp2<- bind_rows(ma.temp2,na.omit(dfMA))

}



```

##Output

```{r data_ou1}

# export DG MARE Med&BS MA table
write.table(format(ma.temp2,digits=3, scientific=F), file=paste("MA.csv",sep=""),
            dec=".",sep=",",col.names=TRUE, row.names=FALSE,na="-1")

head(ma.temp2)
```



