---
title: "Quality check RCG CS format (STREAM project, Task 6.1)"

output:
   word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, warning=FALSE}

library(ggplot2)

# Input to be updated by the user:
setwd("C://Ale//PROGETTI IN CORSO//MARE201622//WP6 STREAM//Task 6.1//DPS9 a priori checks")
name_file=  "DPS9 RCGformat"
immature_stages=c("0","1","2a")
########################################################




data= read.table(paste(name_file,".csv",sep=""),sep=";",header=T)
x=paste("Checking the file ", name_file,".csv",sep="")
```
`r x`

# Check 1: consistency of LFD by year and consistency with allowed ranges

The LFDs available in the sampling data are reported in Figure 1. 

```{r, fig.cap= "Figure 1 - Sampling data: LFDs by year.", fig.width=10, fig.height=8, warning=FALSE,include=TRUE,echo=FALSE}

print(ggplot(data=data, aes(x=Length.class,y= Number.at.length)) + geom_histogram(stat="identity",colour = "blue", fill = "blue", binwidth = 0.5) + facet_grid(Year~ .))

paste("The maximum length class in the data is: ",max(data$Length.class)/10," cm",sep="")
paste("The minimum length class in the data is: ",min(data$Length.class)/10," cm",sep="")

```

If any record resulted outside the allowed range, it is showed below.

```{r, fig.cap= "Figure 1 - Sampling data: LFDs by year.", fig.width=10, fig.height=8, warning=FALSE,include=TRUE,echo=FALSE}

allowed_ranges=read.table("Check_allowed_ranges DPS9.csv",sep=";",header=T)  


error_min_length=data[data$Length.class<allowed_ranges[allowed_ranges$Variable=="Length"]$Min[1], ]
error_max_length=data[data$Length.class>allowed_ranges[allowed_ranges$Variable=="Length"]$Max[1], ]  

if (nrow(error_min_length)!=0 | nrow(error_min_length)!=0){

  print(paste("Records with length outside the allowed range:"), quote=F)  
  print(error_min_length)  
  print(error_max_length)  
  
} else {
print("No error occurred", quote=F)

}
```
# Check 2: consistency of LFD by year and commercial category

The LFDs by commercial category and year are shown in Figure 2.

```{r, fig.cap= "Figure 2 - Sampling data: LFDs by commercial category and by year.", fig.width=10, fig.height=8, warning=FALSE,include=TRUE,echo=FALSE}
# Check consistency of LFD by year and commercial category
ggplot(data=data, aes(x=Length.class,y= Number.at.length)) + geom_histogram(stat="identity",colour = "tan2", fill = "tan2", binwidth = 0.5) + facet_grid(Year~ Commercial.size.category)

```

```{r, fig.cap= "Figure 1 - Sampling data: LFDs by year.", fig.width=10, fig.height=8, warning=FALSE,include=TRUE,echo=FALSE}

# Summary table with ranges by year and commercial category 

pivot_min=aggregate(data$Length.class,by=list(data$Year,data$Commercial.size.category),FUN="min")
colnames(pivot_min)=c("Year","Commercial_size_category","Min")
pivot_max=aggregate(data$Length.class,by=list(data$Year,data$Commercial.size.category),FUN="max")
colnames(pivot_max)=c("Year","Commercial_size_category","Max")
merge(pivot_min,pivot_max, by=c("Year","Commercial_size_category"))

```

# Check 5: Check consistency sex and maturity stage

The scatter plots of length versus maturity stage, by year and by sex are reported in Figure 5. A summary table of the sex and maturity measurements in the sampling data follows.

```{r, fig.cap= "Figure 5 - Sampling data: Scatter plot of maturity stage versus length by year and sex.", fig.width=10, fig.height=8, warning=FALSE,include=TRUE,echo=FALSE}

# Check consistency maturity stage
ggplot(data=data, aes(x=Length.class,y=Maturity.Stage ,col=Sex)) + geom_point(stat="identity", fill = "darkorchid4", binwidth = 0.5) + facet_grid(Year~ Sex)


# summary table of number of individuals by length class by maturity stage
data_sex=data[!is.na(data$Sex),]

tab_sex=aggregate(data_sex$Number.at.length,by=list(data_sex$Year,data_sex$Length.class),FUN="length")
colnames(tab_sex)=c("Year","Length_class","nb_sex_measurements")
tab_sex


# summary table of number of individuals by length class by maturity stage
data_mat=data[!is.na(data$Maturity.Stage),]

tab_mat=aggregate(data_mat$Number.at.length,by=list(data_mat$Year,data_mat$Length.class),FUN="length")
colnames(tab_mat)=c("Year","Length_class","nb_maturity_stage_measurements")
tab_mat

# Maturity ogive
data_mat$Mature[as.character(data_mat$Maturity.Stage) %in% immature_stages]= 0
data_mat$Mature[!(as.character(data_mat$Maturity.Stage) %in% immature_stages)]= 1

mat=data_mat[data_mat$Mature==1,colnames(data_mat) %in%c("Year","Sex","Length.class","Mature","Number.at.length")]
immat=data_mat[data_mat$Mature==0,colnames(data_mat) %in%c("Year","Sex","Length.class","Mature","Number.at.length")]

merge= merge(mat,immat,by=c("Length.class","Year","Sex"),all=TRUE)

colnames(merge)[4]="Mature"
colnames(merge)[6]="Immature"

merge=merge[,c(1,2,3,4,6)]

merge[is.na(merge$Mature),]$Mature=0
merge[is.na(merge$Immature),]$Immature=0
merge$Total=merge$Mature+merge$Immature
```

The maturity ogives estimated by sex are reported in Figure 6 and 7, considering as immature stages `r immature_stages`, aggregating all the maturity data available in the dataset.

```{r, fig.cap= "Figure 6 - Sampling data: Maturity ogive females.", fig.width=5, fig.height=5, warning=FALSE,include=TRUE,echo=FALSE}

#females
merge_temp=merge[as.character(merge$Sex)!="M"& as.character(merge$Sex)!="N",]


years= paste("(",min(merge_temp$Year),"-",max(merge_temp$Year),")",sep="")

Mat=aggregate(merge_temp$Mature,by=list(merge_temp$Length.class),FUN="sum")
Immat=aggregate(merge_temp$Immature,by=list(merge_temp$Length.class),FUN="sum")

merge_temp=merge(Mat,Immat,by=c("Group.1"), all=T)
colnames(merge_temp)=c("Length.class","Mature","Immature")
merge_temp$Total=rowSums(data.frame(merge_temp$Mature,merge_temp$Immature))


mod <- glm(cbind(merge_temp$Mature,merge_temp$Immature) ~ merge_temp$Length.class, family=binomial("logit"))

coeff=coefficients(mod)
L50=-coeff[1]/coeff[2]
IFM=summary(mod)$cov.scaled
I11=IFM[1,1]
I12=IFM[1,2]
I22=IFM[2,2]
SE_L50=sqrt((I11+2*L50*I12+(L50^2)*I22)/(coeff[2]^2))
#inverse of Information Fisher's Matrix
L75=(log(0.75/0.25)-coeff[1])/coeff[2]
L25=(log(0.25/0.75)-coeff[1])/coeff[2]
MR=L75-L25
SE_MR= 2*log(3,exp(1))/((coeff[2])^2)*sqrt(I22) 


plot(merge_temp$Length.class, merge_temp$Mature/merge_temp$Total,cex=1.5,xlab="length(cm)",ylab="proportion of matures",main=paste("Females",years,sep=""))  

lines(merge_temp$Length.class,predict(mod,new.data=merge_temp$Length.class,type="response"),col="deeppink3",lty=1,lwd=3)
legend("bottomright",paste(c("L50= ","MR= "),c(round(L50,3),round(MR,3)),"+/-",c(round(SE_L50,3),round(SE_MR,3))))

```


